Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-04-05 Ben_Manuscripts/hdphmm/Outline.tex
Added some comments.

  %BJC: I'm not sure how much of this I should include. This is pretty much 
  % a reproduction of Fox et al's explanation: https://ieeexplore.ieee.org/abstract/document/5563110
  The infinite-state HMM overcomes this drawback by placing a hierarchical
  Dirichlet process (HDP) prior on the transition probabilities.
  \begin{itemize}
    \item Using some base probability distribution, $H$, a Dirichlet process 
    (DP) generates discrete distributions, $G_0$, over a countably infinite 
    number of probability measures:
    \begin{equation}
      G_0 = \sum_{k=1}^{\infty} \beta_k \delta_{\theta_k} ~~ \theta_k \sim H, \beta \sim GEM(\gamma)
    \end{equation}
    where the $\theta_k$ are values drawn from the base distribution and the
    weights $\beta_k$ come from a stick-breaking process parameterized by the  
    concentration parameter $\gamma$ (equivalently referred to as GEM($\gamma$)). 
    \item Also expressed $G_0 \sim DP(\gamma, H)$
    %MRS3: This next line is a bit vauge (expresses confidence, closely related)
    %BJC3: tried to make more clear
    %MRS4: physically mention what the base distribution is?
   \item The concentration parameter, $\gamma$, expresses one's confidence in the 
    base distribution $H$. 
    \item When $\gamma\to 0$, the first weight of $G_0$, $\beta_1$, approaches 
    unity and for $\gamma\to\infty$, the weights become uniform and $G_0$ closely
    resembles $H$.
    \item Each row, $G_j$, of the transition matrix is produced by drawing from a DP specified 
    using the $\beta$ vector as a discrete base distribution and a separate concentration
    parameter, $\alpha$.
    \begin{equation}
      G_j = \sum_{k=1}^{\infty} \pi_{jk} \delta_{\theta_k} ~~ \pi_j \sim DP(\alpha, \beta)
    \end{equation}
    \item This hierarchical specification ensures that the transition probabilities in 
    each row share the same support points \{$\theta_1$, ..., $\theta_k$\}.
    \item Once the model has converged only a finite number of states will have significant
    sampling.
  \end{itemize}

@@

  %BJC2: I think it would be useful to continue using this trajectory to illustrate 
  % the parameterization process. It might be bunch of plots stacked on each other.
  %BJC2: More labels might be useful if not cluttered
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{prior_guesses.pdf}
  %MRS4: remind people what $c$ physically represents here. 
  \caption{The parameters of the prior on $c$ (black line) should be chosen such
  that the mean levels of each state identified in the trajectory (blue line) lie within
  regions of the prior with reasonable probability. We chose the mean of the prior 
  as halfway between the maximum and minimum of each trajectory dimension. We chose 
  $\sigma$ of the prior by defining the maximum and minimum to be 2 standard deviations
  from the mean (dashed lines).}\label{fig:prior_guesses}
  \end{figure}

@@

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry.
  \begin{itemize}
    %MRS4: Do you need to convert the covariance to cylintrical coordinates here before clustereing? 
    \item We converted the $x$ and $y$ center-of-mass coordinates to $r$.
    \item We fixed the state sequence from the 3D parmeterization and applied the
    inference component of the IHMM procedure to the cylindrical trajectories in
    order to estimate the VAR(1) parameters in terms of $r$ and $z$.
  \end{itemize}

  We clustered like parameter sets in order to reduce the state space to
  a more easily interpretable size.
  \begin{itemize}
  	\item For each solute studied, we identified 200-325 independent states, each
   	with separate VAR(1) parameters.
   	\item Many of these states exhibit very similar dynamical behavior except their
   	mean levels are different, especially in the axial direction where solutes could
   	get trapped along a broad and continuous range of $z$ coordinates. 
  %MRS4: Maybe better to phrase above in terms of the symmetry in z. Why does it matter the range is continuous?
  \end{itemize}

@@

  % BJC2: this part may expand a bit if I try different clustering approaches. But
  % this is the one I'm going with for now.
  We clustered based on the two diagonal entries of $A$ and the two diagonal entries 
  of $\Sigma$.
  \begin{itemize}
   \item One can choose alternative clustering features, such as the eigenvalues of
   the $A$ and $\Sigma$ matrices, but we chose their diagonals because they can be
   easily translated to solute behavior.
   %MRS4: suggest translating them here into solute behavior for the reader!
   \item We also chose to cluster the diagonals of each matrix independently before
   combining the clusters to assign final state labels. 
   \item If we divide the total states into $m$ clusters based on the diagonal entries
   of $A$ and, separately, $n$ clusters based on the diagonal entries of $\Sigma$,
   then there are a total of $mn$ possible cluster combinations.
   \item This allows slightly more flexibility in the total number of states compared
   to clustering on all four features at once, since there can be up to, but not 
   necessarily exactly, $mn$ states. 
   %MRS4: so no clustering based on transition probablity, correct?
  \end{itemize}

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines.
  \begin{itemize}
  	%BJC3: not sure if the commented point worth making.
%    \item We should expect a relatively high number of states. If we only consider
%    the diagonal entries of the $A$ and $\Sigma$ matrices, a total of four independent
%    variables, we can consider clustering each into at least two groups (higher versus
%    lower values), which implies $2^4$ distinct dynamical behaviors. 
    \item We used the Silhouette test in order to score the quality of
    clusters as a function of the number of clusters chosen.
    \item For our data, the Silhouette test generally favors the lowest number
    of clusters possible (see Supporting Information).
    \item However, choosing too few clusters tends to result in 
    inadequate distinguishment between visually obvious differences in 
    dynamic behavior.
    \item This results in finalized parameter sets that are averages of
    distinct behavior which presents further problems with the predictive
    modeling that we discuss later on.
    \item We aimed to maintain the highest Silhouette score, and thus lowest
    number of clusters, possible while verifying that visually distinct 
    states stayed separated.
    \item Based on these guidelines, we decided to group the $A$ and $\Sigma$
    parameters into five clusters each. 
	\item Of the 25 distinct states allowed by this formulation, the solutes
	in this study clustered into 13--18 distinct clusters.
  \end{itemize}

@@

  %BJC3: obviously can use a better thesis.
  We showcase this modeling approach by example. 
  \begin{itemize}
    \item In Figure~\ref{fig:mechanism_map}, we plot a representative 2D methanol 
    trajectory in cylindrical coordinates.
    \item The coordinates are colored according to the cluster with which methanol's
    dynamics are most consistent, which implies that this particular methanol 
    trajctory exhibits 6 distinct dynamical behaviors. 
    \item The solute timeseries are coupled with color coded bars which describe the
    solute's local environment and physical interactions that it undergoes at that
    time frame.
    %MRS: maybe number (1-4) or label (a-d) the rows as well. 
    \item From the top, they describe whether the solute is associated with sodium,
    whether the solute is hydrogen bonding with the monomer as well as what part of
    the monomer it is hydrogen bonding to, the local number density surrounding the
    solute and the radial location of the solute with respect to the nearest pore
    center (this information is redundant but can be a useful visual aid). %BJC3: still debating removing that bar
  \end{itemize}

  %BJC3: I still need labels for this plot
  \begin{figure}
  \includegraphics[width=\textwidth]{mechanism_map.pdf}
  \caption{}\label{fig:mechanism_map}
  \end{figure}

  One can gain significant chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}.
  \begin{itemize}
    \item It is clear that, for methanol, association with sodium is a much rarer
    interaction than hydrogen bonding with the LLC monomers.
    \item Methanol hydrogen bonds to all regions of the monomers dependent on
    its radial position.
    \item Its fluctuations tend to be smallest when hydrogen bonded or in areas
    of high local number density. 
    \item For example, from approximately 75 to 525 ns, methanol appears to be trapped
    in a high density region of the tails.
    \item In the time that follows, it enters a region of relatively low density
    where fluctuations are quite large. 
    \item During this time period, there is intermittent, short-lived hydrogen 
    bonding with the oxygen atoms at the tail ends, as suggested by the frequent 
    state changes during that time period and red slivers in the bar above.
  \end{itemize}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}
  \caption{}\label{fig:dwell_times_MET}
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \end{subfigure}
  \caption{}\label{fig:common_states_MET}
  \end{figure}

@@

  It is important to understand that the clustered parameter sets are
  essentially summaries of many related behaviors. Therefore, the most thorough
  understanding of a specific segment will be gained using descriptive visualizations
  like Figure~\ref{fig:mechanism_map}. But this model very clearly tells the user where
  to look.
  %BJC3: could do a case study of sorts on a chosen segment

  %BJC3: this figure is not even close to done
  %BJC3: I've tried a couple different things to look at the hbonds. The other was 
  % how often they are hydrogen bonded in general versus not hydrogen bond. They all
  % hydrogen bond about the same amount. So I think hbond lifetimes may be the key. 
  % I need to implement this first to test that hypothesis.
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{}\label{fig:hbond_pichart}
  \end{figure}

@@

  % BJC3: It might be good to show a few MD trajectories since they vary somewhat widely
  % in behavior.
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{trajectory_realizations_MET.pdf}
  \caption{}\label{fig:trajectory_realizations_MET}
  \end{figure}

  \noindent The MSDs calculated by stochastic realizations of our model are
  quantitatively similar to MD.
  \begin{itemize}
  	\item For all solutes except acetic acid, our predicted MSD curve lie close to
  	but below the 1$\sigma$ confidence intervals of MD.
  	\item However, in these cases, the average MD MSDs are driven up by single trajectories
    with uncharacteristically large MSDs.
    \item If we remove these uncharacteristic trajectories, the agreement between 
    MD and our realizations improves. 
    % BJC3: I'm not sure the best course of action. I've double checked the bootstrapping
    % procedure and it make sense to me.
    % BJC3: I could show how it's possible to get good agreeement with MD by only
    % generating 24 solute trajectories since in some cases, outliers heavily weight the averages.
    \item Acetic acid has no obvious outliers and shows good agreement between MD and
    IHMM realizations.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_MET.pdf}
  \caption{}\label{fig:msd_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_GCL.pdf}
  \caption{}\label{fig:msd_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_URE.pdf}
  \caption{}\label{fig:msd_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_ACH.pdf}
  \caption{}\label{fig:msd_ACH}
  \end{subfigure}
  \end{figure}

>>>>

  %BJC: I'm not sure how much of this I should include. This is pretty much 
  % a reproduction of Fox et al's explanation: https://ieeexplore.ieee.org/abstract/document/5563110
  The infinite-state HMM overcomes this drawback by placing a hierarchical
  Dirichlet process (HDP) prior on the transition probabilities.
  \begin{itemize}
    \item Using some base probability distribution, $H$, a Dirichlet process 
    (DP) generates discrete distributions, $G_0$, over a countably infinite 
    number of probability measures:
    \begin{equation}
      G_0 = \sum_{k=1}^{\infty} \beta_k \delta_{\theta_k} ~~ \theta_k \sim H, \beta \sim GEM(\gamma)
    \end{equation}
    where the $\theta_k$ are values drawn from the base distribution and the
    weights $\beta_k$ come from a stick-breaking process parameterized by the  
    concentration parameter $\gamma$ (equivalently referred to as GEM($\gamma$)). 
    \item Also expressed $G_0 \sim DP(\gamma, H)$
    \item The concentration parameter, $\gamma$, expresses one's confidence in the 
    base distribution $H$.
    \item We use a uniform base distribution.
    \item When $\gamma\to 0$, the first weight of $G_0$, $\beta_1$, approaches 
    unity and for $\gamma\to\infty$, the weights become uniform and $G_0$ closely
    resembles $H$.
    \item Each row, $G_j$, of the transition matrix is produced by drawing from a DP specified 
    using the $\beta$ vector as a discrete base distribution and a separate concentration
    parameter, $\alpha$.
    \begin{equation}
      G_j = \sum_{k=1}^{\infty} \pi_{jk} \delta_{\theta_k} ~~ \pi_j \sim DP(\alpha, \beta)
    \end{equation}
    \item This hierarchical specification ensures that the transition probabilities in 
    each row share the same support points \{$\theta_1$, ..., $\theta_k$\}.
    \item Once the model has converged only a finite number of states will have significant
    sampling.
  \end{itemize}

@@

  %BJC2: I think it would be useful to continue using this trajectory to illustrate 
  % the parameterization process. It might be bunch of plots stacked on each other.
  %BJC2: More labels might be useful if not cluttered
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{prior_guesses.pdf}
  %MRS4: remind people what $c$ physically represents here. 
  \caption{The parameters of the prior on the mean vector, $c$, (black line) should be chosen such
  that the mean levels of each state identified in the trajectory (blue line) lie within
  regions of the prior with reasonable probability. We chose the mean of the prior 
  as halfway between the maximum and minimum of each trajectory dimension. We chose 
  $\sigma$ of the prior by defining the maximum and minimum to be 2 standard deviations
  from the mean (dashed lines).}\label{fig:prior_guesses}
  \end{figure}

@@

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry.
  \begin{itemize}
    %MRS4: Do you need to convert the covariance to cylintrical coordinates here before clustereing? 
    %BJC4: No because I only convert the coordinates to cylindrical and then use 
    % the inference part of the IHMM to estimate the covariance.
    \item We converted the $x$ and $y$ center-of-mass coordinates to $r$.
    \item We fixed the state sequence from the 3D parmeterization and applied the
    inference component of the IHMM procedure to the cylindrical trajectories in
    order to estimate the VAR(1) parameters in terms of $r$ and $z$.
  \end{itemize}

  We clustered like parameter sets in order to reduce the state space to
  a more easily interpretable size.
  \begin{itemize}
  	\item For each solute studied, we identified 200-325 independent states, each
   	with separate VAR(1) parameters.
   	\item Many of these states exhibit very similar dynamical behavior except their
   	mean levels are different, especially in the axial direction where solutes could
   	gßet trapped along a broad and continuous range of $z$ coordinates. 
  %MRS4: Maybe better to phrase above in terms of the symmetry in z. Why does it matter the range is continuous?
  \end{itemize}

@@

  % BJC2: this part may expand a bit if I try different clustering approaches. But
  % this is the one I'm going with for now.
  We clustered based on the two diagonal entries of $A$ and the two diagonal entries 
  of $\Sigma$.
  \begin{itemize}
   \item One can choose alternative clustering features, such as the eigenvalues of
   the $A$ and $\Sigma$ matrices, but we chose their diagonals because they can be
   easily translated to solute behavior.
   %MRS4: suggest translating them here into solute behavior for the reader!
   \item We also chose to cluster the diagonals of each matrix independently before
   combining the clusters to assign final state labels. 
   \item If we divide the total states into $m$ clusters based on the diagonal entries
   of $A$ and, separately, $n$ clusters based on the diagonal entries of $\Sigma$,
   then there are a total of $mn$ possible cluster combinations.
   \item This allows slightly more flexibility in the total number of states compared
   to clustering on all four features at once, since there can be up to, but not 
   necessarily exactly, $mn$ states. 
   %MRS4: so no clustering based on transition probablity, correct?
   %BJC4: not at the moment. That might have been more helpful if I was first-order differencing.
  \end{itemize}

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines.
  \begin{itemize}
  	%BJC3: not sure if the commented point worth making.
%    \item We should expect a relatively high number of states. If we only consider
%    the diagonal entries of the $A$ and $\Sigma$ matrices, a total of four independent
%    variables, we can consider clustering each into at least two groups (higher versus
%    lower values), which implies $2^4$ distinct dynamical behaviors. 
%MRS5: should be lower-case 's' on silhouette
    \item We used the Silhouette test in order to score the quality of
    clusters as a function of the number of clusters chosen.
    \item For our data, the Silhouette test generally favors the lowest number
    of clusters possible (see Supporting Information).
    \item However, choosing too few clusters tends to result in 
    inadequate distinguishment between visually obvious differences in 
    dynamic behavior.
    \item This results in finalized parameter sets that are averages of
    distinct behavior which presents further problems with the predictive
    modeling that we discuss later on.
    \item We aimed to maintain the highest Silhouette score, and thus lowest
    number of clusters, possible while verifying that visually distinct 
    states stayed separated.
    \item Based on these guidelines, we decided to group the $A$ and $\Sigma$
    parameters into five clusters each. 
	\item Of the 25 distinct states allowed by this formulation, the solutes
	in this study clustered into 13--18 distinct clusters.
  \end{itemize}

@@

  %BJC3: obviously can use a better thesis.
  We showcase this modeling approach by example. 
  \begin{itemize}
    \item In Figure~\ref{fig:mechanism_map}, we plot a representative 2D methanol 
    trajectory in cylindrical coordinates.
    \item The coordinates are colored according to the cluster with which methanol's
    dynamics are most consistent, which implies that this particular methanol 
    trajctory exhibits 6 distinct dynamical behaviors. 
    \item The solute timeseries are coupled with color coded bars which describe the
    solute's local environment and physical interactions that it undergoes at that
    time frame.
    %MRS: maybe number (1-4) or label (a-d) the rows as well. 
    %BJC4: yeah, meant to do that. As it stands, I'm referring to 1 through 4 as bottom through top.
    \item From the top, they describe whether the solute is associated with sodium,
    whether the solute is hydrogen bonding with the monomer as well as what part of
    the monomer it is hydrogen bonding to, the local number density surrounding the
    solute and the radial location of the solute with respect to the nearest pore
    center (this information is redundant but can be a useful visual aid). %BJC3: still debating removing that bar
  \end{itemize}

  %BJC3: I still need labels for this plot
  %BJC4: Also, for some reason, when saving to a PDF, the colorbars are faded. I'll fix this eventually.
  %MRS5: explicitly state how many states there are, and their colors.
  \begin{figure}
  \includegraphics[width=\textwidth]{mechanism_map.pdf}
  \caption{We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a 
  %MRS5:
  %chosen
  single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters %
  %MRS5:
  %are represented by 
  place them in
  the same 
  %BJC4: eventually I'll label the bars and probably put a picture of the monomer with different oxygens labeled
  cluster. 
  %MRS5:
  %Above 
  On the top of
the plot are bars which represent different physical interactions and
  the membrane's local environment as a function of time. The top bar is colored green 
  at time points where solutes are associated with sodium. Methanol associates
  with sodium ions relatively rarely. The next bar down is colored when hydrogen bonds exist. Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  corrrespond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol is hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. The bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  %BJC4: eventually I want some kind of color bar legend
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion appears restricted (e.g. from 75 to 550 ns). It is low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). Finally, the bottom bar
  simply color-codes the solute's radial distance from the pore center. This information
  is the same as shown by the radial coordinate but can be a helpful visual aid when
  studying the other bars.}\label{fig:mechanism_map}
  \end{figure}
  %MRS5: hmm, one sign that someting might be nonideal is the clustering in the light blue state (assuming 1 light blue state), looking at the 75 nm to 550 nm segment and the 900 - 1050 nm segment, the r is quite different (1 vs 3), and the h-bonding patterns are very different (Green vs. none). So it's not entirely clear these should necessarily be the same state. Thoughts?
  One can gain significant chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}.
  \begin{itemize}
    \item It is clear that, for methanol, association with sodium is a much rarer
    interaction than hydrogen bonding with the LLC monomers.
    \item Methanol hydrogen bonds to all regions of the monomers dependent on
    its radial position.
    \item Its fluctuations tend to be smallest when hydrogen bonded or in areas
    of high local number density. 
    \item For example, from approximately 75 to 525 ns, methanol appears to be trapped
    in a high density region of the tails.
    \item In the time that follows, it enters a region of relatively low density
    where fluctuations are quite large. 
    \item During this time period, there is intermittent, short-lived hydrogen 
    bonding with the oxygen atoms at the tail ends, as suggested by the frequent 
    state changes during that time period and red slivers in the bar above.
  \end{itemize}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some errorbars on this by drawing the transition probabilities 
  % from a dirichlet distribution.
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}
  \caption{}\label{fig:dwell_times_MET}
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to the majority of methanol trajectories. In (a), we show representative
  dynamics of four states that are common to more than half of the solute trajectories. The
  percentage of trajectories in which they appear are listed between the $r$ and $z$ plots. 
  All four of these states are exhibited in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  %MRS5: state how much shifted - or maybe plot a dotted line at the zero of each plot?
  shifted for the purpose of visualizing them. (b) Using the transition probabilities (p), 
  we estimated the average time spent within each of the states using 1000 simulations of a Bernoulli 
  process where we continued to draw until transitioning out of the state with probability 1 - p.
  On average, methanol stay in states 2 and 4 for much longer periods of time than states 1 and 3.
  This is qualitatively evident in Figure~\ref{fig:mechanism_map}. (c) We can use the understanding
  from (a) and (b) in order to relate the state parameters to solute motion. States 1 and 3 have
  large covariances relative to states 2 and 4. Paired with their short dwell times, it is reasonable
  to hypothesize that states 1 and 3 contribute to hopping behavior while states 2 and 4 
  correspond to trapped solute behavior. All states have positive autoregressive coefficients (A) in 
  each dimension which implies motion that is positively correlated to its previous step. States 3 and
  4 have higher autoregressive coefficients than 1 and 2. This implies that solutes in states 3 and 
  4 tend to wander more from their mean position in each dimension. This appears to be the case when
  comparing the behavior of states 2 and 4 in Figure~\ref{fig:mechanism_map}.}\label{fig:common_states_MET}
  \end{figure}

@@

  It is important to understand that the clustered parameter sets are
  %MRS5: summaries-> averages?
  essentially summaries of many related behaviors. Therefore, the most thorough
  %MRS5->Figure 6 does have show that there is variance between trajectories. 
  understanding of a specific segment will be gained using descriptive visualizations
  like Figure~\ref{fig:mechanism_map}. But this model very clearly tells the user where
  to look.
  %BJC3: could do a case study of sorts on a chosen segment

  %BJC3: this figure is not even close to done
  %BJC3: I've tried a couple different things to look at the hbonds. The other was 
  % how often they are hydrogen bonded in general versus not hydrogen bond. They all
  % hydrogen bond about the same amount. So I think hbond lifetimes may be the key. 
  % I need to implement this first to test that hypothesis.
  %BJC4: I didn't put sodium ion assocaition in this because it's infrequent. 
  %MRS5: is it correlated with specific states?
  %BJC4: If I compare to acetic acid or urea, I could add it back. I was thinking of splitting
  % the pi chart in half with slices corresponding to hbonds on one side and association on the other
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{With our clustering approach, methanol exhibits similar state behaviors
  %MRS5: So this is a little counterintuitive, in that the separate states have similar behavior.
  %MRS5: Are there charts that show what is _different_ about the states as well?  The previous figure
  %      Shows that they are dynamically different, but this structural pie charge says the existance
  %      of H-bonding is not the reason . . . 
  in all regions of the membrane. In all of the common states described in 
  Figure~\ref{fig:common_states_MET}, methanol, at some point, hydrogen bonds with 
  each region of the LLC monomer to an appreciable degree. However, methanol frequently
  is not hydrogen bonded at all.
  %BJC4: hypothesis yet to be tested. Somehow, I'll incorporate lifetimes into this plot
  %MRS4: lifetimes would be good!
  We can learn more by looking at the hydrogen bond lifetime. In states 2 and 4, solutes
  stay hydrogen bonded for longer periods of time.
  %MRS5: OK, here is some physical insight - local density is what determines trapping. Emphasize the physical mechanisms (wat does and doesn't come out physically because of these analyses) like this more explicitly.
  %MRS5: can you make other graphs like the H-bond graph? like histogram/pie chart of local density
  %MRS5: as a function of state?  What about r as a function of state?  what other ways are there to show
  %MRS5: the characteristics of the states?
  These longer lifetimes exist because, when hydrogen bonded, the local density is higher, preventing the 
  bond from breaking so easily. In states 1 and 3, hbonds are fleeting interactions because
  the surrounding density is low, allowing solutes to move more freely.
  }\label{fig:hbond_pichart}
  \end{figure}

@@

  % BJC3: It might be good to show a few MD trajectories since they vary somewhat widely
  % in behavior.
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{trajectory_realizations_MET.pdf}
  %MRS5: maybe it would be good to put a dotted line through each of the centers of the offset trajectories?
  %MRS5: Do they really qualitatively reproduce?  the r behavior doesn't really look that similar, as it goes back
  %      and forth, but not get stuck. 
  %      Note: it's OK that if the model ends up not being able to describe the system, but we should be clear in 
  %      what ways it does and doesn't.

  \caption{We can qualitatively reproduce the average behavior of methanol trajectories
  using our model. Solute trajectories generated by our model (blue) show the same
  hopping and trapping behavior exhibited by MD (black). The behavior of individual 
  MD trajectories tend to show wider variability than our model's realization because
  the model effectively represents the average behavior of the MD trajectories. This 
  implies that much longer simulations might be necessary in order to obtain a set
  of trajectories that explore state space to the same extent.
  %BJC4: this observation is how I got the idea to calculate MSD by parameterizing
  % individual trajectories  
  %BJC4: Also note that I'm not currently happy with how I generate the radial dimension. 
  %MRS5: how do you do that currently? We talked about generating dynamics in x, y, and z, 
  %MRS5: (should be same parameters for x and y, because of symmetry), and then converting to r. 
  %MRS5: could make this more clear in the methods. 
  % Improvements will not affect z or the MSD because the VAR parameters won't change except for c
   }\label{fig:trajectory_realizations_MET}
  \end{figure}

  \noindent The MSDs calculated by stochastic realizations of our model are
  quantitatively similar to MD.
  \begin{itemize}
  	\item For all solutes except acetic acid, our predicted MSD curve lie close to
  	but below the 1$\sigma$ confidence intervals of MD.
  	\item However, in these cases, the average MD MSDs are driven up by single trajectories
    with uncharacteristically large MSDs.
    \item If we remove these uncharacteristic trajectories, the agreement between 
    MD and our realizations improves. 
    % BJC3: I'm not sure the best course of action. I've double checked the bootstrapping
    % procedure and it make sense to me.
    % MRS3: Thinking out loud - if 4/24 trajectories are big, then 
    % (5/6)^24 = 0.013 will contain none of the 'big' trajectories.
    % 24*(1/6)*(5/6)^23 = 0.060 will contain 1.
    % C(24,2)*(1/6)^2*(5/6)^22 = 0.139 will contain 2
    % C(24,3)*(1/6)^3*(5/6)^21 = 0.204 will contain 3 
    % C(24,4)*(1/6)^4*(5/6)^20 = 0.214 will contain 4
    % C(24,5)*(1/6)^5*(5/6)^19 = 0.171 will contain 5
    % C(24,6)*(1/6)^6*(5/6)^18 = 0.108 will contain 6
    % MRS5: so I guess it makes sense that the median bootstrap trajectory will have 4 big trajectories,
    % MRS5: and that the 1 sigma line on the lower end will be ones with 2 'big' trajectories.
    % MRS5: You could also plot the 95% confidence intervals
    % BJC3: I could show how it's possible to get good agreeement with MD by only
    % generating 24 solute trajectories since in some cases, outliers heavily weight the averages.
    \item Acetic acid has no obvious outliers and shows good agreement between MD and
    IHMM realizations.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_MET.pdf}
  \caption{}\label{fig:msd_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_GCL.pdf}
  \caption{}\label{fig:msd_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_URE.pdf}
  \caption{}\label{fig:msd_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_ACH.pdf}
  \caption{}\label{fig:msd_ACH}
  \end{subfigure}
  \caption{The MSDs predicted based on 1000 realizations of our model (blue) tend to underpredict
  the MSDs based on the MD trajectories (black). Due to the variability in solute behavior across MD trajectories,
  as demonstrated in Figure~\ref{fig:trajectory_realizations_MET}, there are several 
  individual solute MSD curves that are significantly higher than the rest and drive up the
  mean of the MSD. If we remove those trajectories from the MSD calculation, we obtain 
  estimates much closer to the IHMM estimate.(I haven't done it yet, but could plot a third curve
  with outlier removed). Acetic acid, has no obvious outliers which is consistent with our model's
  prediction which lies within the 1$\sigma$ confidence interval of MD after a 1000 ns time lag.}\label{fig:msds}
  \end{figure}
